---
title: "Financial Analysis-PCA"
author: ""
date: "11/25/2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
## Introduction
In This part of the project, a Principal component analysis is to be conducted on the dataset which consists of the following:
Three sectors and five representative firm of each sector, selected macroeconomic indicators and Fama-French factors.

```{r getdata, include=FALSE}
options(warn=-1)

## install packages are hidden, just delete ## to install the packages
library(dplyr)
##install.packages('factoextra')
library(factoextra);
library(tidyr);
library(reshape2);
library(readxl)
##install.packages('FactoMineR')
library(FactoMineR)
##install.packages('PerformanceAnalytics')
library(PerformanceAnalytics)
require(quantmod)
##install.packages('corrplot')
library(corrplot)
## in this part we will load the data using the following function 
read_excel_allsheets <- function(filename, tibble = FALSE) {
    sheets <- readxl::excel_sheets(filename)
    x <- lapply(sheets, function(X) readxl::read_excel(filename, sheet = X))
    if(!tibble) x <- lapply(x, as.data.frame)
    names(x) <- sheets
    x
}
## this should be changed according to where the data files are located in your machine
filepath1<- "~/downloads/MonthlyMacro_20171018.xlsx"
filepath2 <- "~/downloads/monthlydata.xlsx"
## The warnings can be disregarded
macroFact <- read_excel_allsheets(filepath1)
companyData <- readxl::read_excel(filepath2, col_types = c("date","numeric","numeric","numeric","numeric","numeric","numeric","numeric","numeric","numeric","numeric","numeric","numeric","numeric","numeric","numeric","numeric","numeric","numeric","numeric"),skip =3);
companyData <- companyData[-c(1, 2), ]
names(companyData)[1] <- 'Year'

```



```{r datafix, include=FALSE}
companyData[1] <- lapply(companyData[1], function(x) format(x,'%Y-%m' ))
companyData[1]<- lapply(companyData[1], function(x) as.Date(paste0(as.character(x), '-01'), format='%Y-%m-%d' ))
companylogData<- tbl_df(Return.calculate(xts(companyData[, -1], order.by=as.POSIXct(companyData$Year)), method = "log"))
companylogData['Year']<- sort(companyData$Year)
companyData2 <- tbl_df(Return.calculate(xts(companyData[, -1], order.by=as.POSIXct(companyData$Year)), method = "discrete"))
companyData2['Year']<- sort(companyData$Year)
companylogData<- companylogData[complete.cases(companylogData), ]
companyData2<- companyData2[complete.cases(companyData2), ]
sources<- macroFact[2]
sources <- sources[[1]]
frenchfam <- sources[][-c(1, 2),]
names(frenchfam) <- c('Year' ,frenchfam[1 , 2:5])
frenchfam <- frenchfam[-1, ]
frenchfam[1] <- lapply(frenchfam[1], function(x) as.Date(paste0(as.character(x), '01'), format='%Y%m%d'))
frenchfam <- frenchfam[complete.cases(frenchfam), ]

sources2<- macroFact[3]
sources2<- sources2[[1]][-c(1, 2),]
emp <- sources2[-c(1:9),]
names(emp)<- emp[1,]
emp <- emp[-1, ]
emp <- melt(emp, id = 'Year')
emp <- unite(emp, 'Year' ,c('Year','variable'), sep = "", remove = TRUE)
names(emp)<- c('Year', 'emp_value')
emp[1]<- lapply(emp[1], function(x) as.Date(paste0(as.character(x), '01'), format='%Y %B%d' ))
emp <- emp[complete.cases(emp), ]


sources3<- macroFact[4]
sources3<- sources3[[1]][-c(1, 2),]
cpi <- sources3[-c(1:8),]
names(cpi)<- cpi[1,]
cpi <- cpi[-1, ]
cpi <- melt(cpi, id = 'Year')
cpi <- unite(cpi, 'Year' ,c('Year','variable'), sep = "", remove = TRUE)
names(cpi)<- c('Year', 'cpi_value')
cpi[1]<- lapply(cpi[1], function(x) as.Date(paste0(as.character(x), '01'), format='%Y %B%d' ))
cpi <- cpi[complete.cases(cpi), ]

sources4<- macroFact[6]
sources4<- sources4[[1]][-c(1, 2),]
ppi <- sources4[-c(1:6),]
names(ppi)<- ppi[1,]
ppi <- ppi[-1, ]
ppi <- melt(ppi, id = 'Year')
ppi <- unite(ppi, 'Year' ,c('Year','variable'), sep = "", remove = TRUE)
names(ppi)<- c('Year', 'ppi_value')
ppi[1]<- lapply(ppi[1], function(x) as.Date(paste0(as.character(x), '01'), format='%Y %B%d' ))
ppi <- ppi[complete.cases(ppi), ]


sources5<- macroFact[7]
sources5<- sources5[[1]][-c(1, 2),]
import <- sources5[-c(1:7),]
names(import)<- import[1,]
import <- import[-1, ]
import <- melt(import, id = 'Year')
import <- unite(import, 'Year' ,c('Year','variable'), sep = "", remove = TRUE)
names(import)<- c('Year', 'import_value')
import[1]<- lapply(import[1], function(x) as.Date(paste0(as.character(x), '01'), format='%Y %B%d' ))
import <- import[complete.cases(import), ]


sources6<- macroFact[8]
sources6<- sources6[[1]][-c(1, 2),]
xport <- sources6[-c(1:7),]
names(xport)<- xport[1,]
xport <- xport[-1, ]
xport <- melt(xport, id = 'Year')
xport <- unite(xport, 'Year' ,c('Year','variable'), sep = "", remove = TRUE)
names(xport)<- c('Year', 'xport_value')
xport[1]<- lapply(xport[1], function(x) as.Date(paste0(as.character(x), '01'), format='%Y %B%d' ))
xport <- xport[complete.cases(xport), ]

sources7<- macroFact[5]
sources7<- sources7[[1]][-c(1, 2),]
manuf <- sources7[-c(1:8),]
manuf <- manuf[-1,-c(1,2)]
names(manuf)<- c('Year', 'manuf_value')
manuf[1]<- lapply(manuf[1], function(x) as.Date(paste0(as.character(x), '01'), format='%Y %B%d' ))
manuf <- manuf[complete.cases(manuf), ]


all_data_simple<- Reduce(function(x, y) merge(x, y, all=FALSE, by= 'Year'), list(companyData2, emp, xport,import,ppi,cpi,manuf,frenchfam))
all_data_log <- Reduce(function(x, y) merge(x, y, all=FALSE, by= 'Year'), list(companylogData, emp, xport,import,ppi,cpi,manuf,frenchfam))
```

## An overlook of the Data

There are two datasets
1. a simple return 
2. log return

```{r dataover, echo = FALSE, tidy= TRUE}
all_data_simple[,-1]<- lapply(all_data_simple[,-1], function(x) as.numeric(x))
all_data_log[,-1]<- lapply(all_data_log[,-1], function(x) as.numeric(x))

head(all_data_simple)
head(all_data_log)
```


## The analysis

After conducting the analysis we have the following results

```{r analys, echo = FALSE, tidy = TRUE}
myPCAsimple <- PCA(all_data_simple[-1], scale.unit = TRUE, graph = FALSE)

myPCAlog <- PCA(all_data_log[-1], scale.unit = TRUE, graph = FALSE)

simpleigenvalue <- get_eigenvalue(myPCAsimple)

logeigenvalue <- get_eigenvalue(myPCAlog)
simpleigenvalue

logeigenvalue
```

## Figures and plots

```{r plotanalys, tidy= TRUE}
fviz_eig(myPCAsimple, addlabels = TRUE, ylim = c(0, 50), title= 'Scree Plot of Principle Components'
                        )
```
```{r plotanaly2, tidy= TRUE}




fviz_pca_var(myPCAsimple, col.var = "cos2",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), 
             repel = TRUE # Avoid text overlapping
             ,title= 'Biplot of PC1 and PC2'
                        
             )
            

             

```



```{r plotanaly3, tidy= TRUE, fig.width= 20, fig.height=12 }


corrplot(myPCAsimple$var$cos2, is.corr=FALSE)

```
The correlation between each variable and PC 
\newpage
## Analysis


From the Exploratory analysis above, we can view how each variable contribute to each PC
In order for us to represent certain percentage of the dataset, we can view the percentage of each Principle component.
to represent:
 50%  we need PC1-PC3
 75% we need PC1-PC6
 90% we need PC1-PC13